---
title: "Colorado"
author: "Marcela Suárez"
date: "July 25, 2019"
output: html_document
---
<br>


This document presents the results obtained for Colorado.  

<br>

## **Analysis with the whole dataset**

#### **A general  overview of the dataset**

```{r import_data, echo=FALSE, message=FALSE}

library(tidyverse) 
colorado <- read_csv("ColoradoFloodOriginal.csv") # Loading data into R

 # Create a subset with the variables to be used
library(lubridate) # For dates
colorado <- colorado %>% 
  mutate(date = ymd_hms(colorado$created_at, tz='America/Denver')) %>% # Mountain Daylight Time. Time zone in Colorado (GMT-6)
  select(lat = latitude, 
         lon = longitude, 
         date = date, 
         state = c_state,
         text = t_text,
         user = u_name,
         u_profile = u_description)

str(colorado) # shows the structure of the dataframe
```

Tweets were collected from `r min(colorado$date)` to `r max(colorado$date)`  

<br>


####  **Temporal stages**

Four stages were defined:

```{r define_temporal_stages, echo=FALSE, message=FALSE}

colorado <- colorado %>%  # The way to define them in a tidyway
  mutate(stage = ifelse(date <= "2013-09-11 00:00:00 MDT", "preflood", 
                        ifelse(date <= "2013-09-16 00:00:00 MDT", "flood",
                               ifelse(date <= "2013-09-23 00:00:00 MDT", "immediate_aftermath", "postflood")))) %>%
  mutate(stage = factor(stage, levels = unique(stage))) %>% # Turning stages into factors to use them for plotting later
  mutate(stage = factor(stage,levels(stage)[c(3, 4, 2, 1)])) # To plot in an specific order

```

* *Preflood*: From `r min(colorado$date)` to 2013-09-11 00:00:00.
* *Flood*: From 2013-09-11 00:00:00 to 2013-09-16 00:00:00.
* *Immediate Aftermath*: From 2013-09-16 00:00:00 to 2013-09-23 00:00:00.
* *Postflood*: From 2013-09-23 00:00:00 to `r max(colorado$date)`.

<br>


####  **Cleaning dataset**

* *Removing retweets:*
```{r remove_retweets, echo=FALSE, message=FALSE}

library(tidytext)
library(stringr)

colorado <- colorado %>% 
  filter(!str_detect(text, "^RT")) %>% 
  filter(!str_detect(text, "#TweetMyJobs")) %>% 
  filter(!str_detect(text, "#Job")) %>% 
  filter(!str_detect(text, "I'm")) # There will remain some retuits since those will be commented retweets (actually adding some info)
```
    From the original dataset containing 5915 tweets, only 5846 were retained.  

* *Removing tweets sent by bots:*
```{r remove_bots, echo=FALSE, message=FALSE}
colorado <- colorado %>% 
  filter(!str_detect(u_profile, "TwitZip")) %>% 
  filter(!str_detect(u_profile, "TMJ-COB PM Jobs")) %>% 
  filter(!str_detect(u_profile, "JobsDirectUSA.com")) %>% 
  filter(!str_detect(u_profile, "Whole Foods"))

```
    Tweets sent by [TwitZIP](http://twitzip.com/), a bot service to discover local neighborhood news, were removed. After this process 4916 were retained. Other three bot accounts *("TMJ-COB PM Jobs", "JobsDirectUSA.com", "Whole Foods")* were detected and tweets were removed. A total of 4158 tweets were retained.
    
* *Removing urls, usernames, punctuation and numbers:*
```{r clean_tweets, echo=FALSE, message=FALSE}

colo_clean <- colorado %>% 
   mutate(text = str_remove_all(text, 'http[^ ]+'), #Finally removing urls in a nice way
          text = str_remove_all(text, '@[^ ]+'), # Removing usernames
          text = str_remove_all(text, "[:punct:]"), # Removing puntuation
          text = str_remove_all(text, "[:digit:]"), # Removing numbers
          text = str_remove_all(text, "|")) 
```
    All references to usernames, urls and punctuation and numbers were removed from each tweet in order to prepare the dataset for tokenization.

<br>



####  **Most common words**

A quick view of the most common words in the whole dataset: 

```{r one-token-per-document-per-row, echo=FALSE, message=FALSE}

colo_tidy <- colo_clean %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) # remove stopwords

colo_tidy %>% 
  count(word, sort = TRUE) # Count words
```

Since **"boulder"** is the most common word and is going ot have a big effect in our topic modelling, it was removed from the dataset. The following five terms *("boulderflood", "colorado", "coflood", "cowx", "flood")* were also excluded because they were so common and used neutrally in all four stages. After excluding these six terms, the new list of common words looks as follows:
                                 
                                 
```{r remove_words, echo=FALSE, message=FALSE}

keywords <- c("boulder", "boulderflood", "colorado", "coflood", "cowx", "flood")
colo_tidy <- colo_clean %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) %>%  # remove stopwords
  filter(!word %in% keywords) # Next two candidates to be removed are Im and boulderflood. TO CHECK LATER!

colo_tidy %>% 
  count(word, sort = TRUE) # Count words
```

<br>

The statistic **tf-idf** was computed in order to identify which words are important in each of the flood stages. It measures how important a word is to a document in a collection (or corpus) of documents, in our case, it measures how important a word is to a tweet in a collection of tweets. In this case the collection of tweets was the set of tweets belonging to each stage.

```{r explore_tf_idf, echo=FALSE, message=FALSE}

colo_tf_idf <- colo_tidy %>% 
  count(stage, word, sort = TRUE) %>% 
  bind_tf_idf(word, stage, n) %>%
  arrange(-tf_idf) %>%
  group_by(stage) %>% 
  top_n(10) %>% 
  ungroup 

colo_tf_idf %>% 
  mutate(word = reorder_within(word, tf_idf, stage)) %>% #reordering by tf_idf
  ggplot(aes(word, tf_idf, fill = stage)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~stage, scales = "free", ncol = 2)+
  scale_x_reordered() +
  coord_flip() +
  theme(strip.text=element_text(size=11)) +
  labs(x = NULL, y = "tf-idf",
       title = "Highest tf-idf words in tweets during the different stages of Colorado Flood",
       subtitle = "Words importance within tweets sent during each stage")


```

<br>



####  **Topic Modeling**

After playing with different numbers, I decided to train a topic model with 4 topics. From 5 on, topics started to look very similar (with the same bag of words). Here is a summary of the results after this process: 

```{r topic_modeling, echo=FALSE, message=FALSE}
library(stm) #fast compared to other implementations of topic models. Base don C++
library(quanteda)

colo_dfm <- colo_tidy %>%
  count(stage, word, sort = TRUE) %>% 
  cast_dfm(stage, word, n) # 4 documents (4 stages). Quanteda document frecuency. Special implementation for document term matrix

# Let's train oir topic model

topic_model <- stm(colo_dfm, K = 4, init.type = "Spectral") # unsupervised 
summary(topic_model)
```

The probabilities that each word is generated from each topic are shown below:

```{r beta, echo=FALSE, message=FALSE}
td_beta <- tidy(topic_model) # The beta matrix shows what are the words that contribute to each topic.

td_beta %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  ungroup %>% 
  mutate(topic = paste0("Topic ", topic),
         term = reorder(term, beta)) %>% #reordering by tf_idf
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~topic, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics") # To see which words contribute the most to each topic
```

From this, we can't demonstrate yet a clear association of these topics with the four stages of the disaster.

On the other hand, besides estimating each topic as a mixture of words, LDA also models each document (tweet in our case) as a mixture of topics. We can examine the per-document-per-topic probabilities, called the "gamma probability". So, looking at the probability that each document (tweet) is generated from each topic:


```{r gamma, echo=FALSE, message=FALSE}
td_gamma <- tidy(topic_model, matrix = "gamma",                    
                 document_names = rownames(colo_dfm))

ggplot(td_gamma, aes(gamma, fill = as.factor(topic))) +
  geom_histogram(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~ topic, ncol = 2) +
  labs(title = "Distribution of document probabilities for each topic",
       subtitle = "Each topic is associated with 1-3 stages",
       y = "Number of stages", x = expression(gamma))

```

From the plot above, we can see that each stage is strongly associated with a single topic. Gamma probabilities tell us which topics are coming from which documents (tweets).

<br>


## **Analysis with a subset of the data**

Now I will repeat the same analysis but considering tweets sent from what should be the "affected area".

<br>

####  **Spatial Clustering**

In order to identify tweets sent from the afected area, a hierarchical spatial clustering was applied to the data.

Before doing so, the data was first projected in North America Lambert Conformal Conic.

```{r getting_spatial, echo=FALSE, message=FALSE}

library(sf)

# Store tweets as simple features and project data
colo_sf_p <- colo_clean %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)


colo_sf <- colo_sf_p %>% # set WGS84 as original datum
  st_transform(crs ="+proj=lcc +lat_1=20 + lat_2=60 + lat_0=40 +
               lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +
               units=m no_defs") # Projected in North_America_Lambert_Conformal_Conic

st_crs(colo_sf) # Retrieve current coord ref system: EPSG: 4326 WGS84

```

Then, a hierarchical implementation of dbscan was used **(hdscan)**, considering a minimum of 350 tweets per cluster.

```{r spatial_clusters, echo=FALSE, message=FALSE}

library(dbscan)
library(leaflet)
library(htmltools)


set.seed(123)

clusters <- hdbscan(colo_sf %>%
                      st_coordinates(), #This rounds coordinates
                    minPts = 550)

colo_clusters <- colo_sf %>% 
  mutate(cluster = clusters$cluster)

## Plotting spatial cluster results

colo_clusters <- colo_clusters %>% # Need to reproject in WGS84 datum. long lat format.
  st_transform(crs = 4326)

colo_clusters$cluster <- as.factor(colo_clusters$cluster) #Clusters as factors for coloring
pal <- colorFactor(c("#636363", "red", "Blue"), domain = c("0", "1", "2"))

# #1. or Red cluster has 608 tweets. Denver
# #2. or Blue cluster has 1998 tweets. Boulder
# 2219/4840 tweets were classified as outliers.
```


```{r map_clusters, echo=FALSE, message=FALSE}

coloMap <- leaflet(colo_clusters) %>% # Interactive map to see resulting clusters
  addTiles()  %>%
  addProviderTiles(providers$OpenStreetMap.BlackAndWhite) %>% 
  addCircles(weight = 3, 
             radius=40,
             color= ~pal(cluster), 
             stroke = TRUE, 
             fillOpacity = 0.5,
             popup = ~htmlEscape(cluster))%>% 
  setView(lng = -105, lat = 40, zoom = 8.5)

coloMap

```



####  **Most common words (Within Spatial Clusters)**

A quick view of the most common words in the whole dataset: 

```{r data_subsets_tidy, echo=FALSE, message=FALSE}

#This is mainly to remove geometry to make computation faster

clusters_df <- colo_clusters %>% 
  filter(cluster == 1 | cluster == 2) %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  st_set_geometry(NULL)

```


```{r one-token-per-document-per-row-clusters, echo=FALSE, message=FALSE}

clusters_tidy <- clusters_df %>% 
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words)  # remove stopwords
  
clusters_tidy %>% 
  count(word, sort = TRUE) # Count words

```


Again, since **"boulder"** is the most common word and is going to have a big effect in our topic modelling, it was removed from the dataset. The following four terms *("boulderflood", "colorado", "cowx", "flood")* were also excluded because they were so common and used neutrally in all four stages. After excluding these six terms, the new list of common words looks as follows:
                                 
                                 
```{r remove_words_clusters, echo=FALSE, message=FALSE}

keywords_clusters <- c("boulder", "boulderflood", "colorado", "cowx", "flood")

clusters_tidy <- clusters_df %>% 
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) %>%   # remove stopwords
  filter(!word %in% keywords_clusters) 

clusters_tidy %>% 
  count(word, sort = TRUE) # Count words

```


```{r explore_tf_idf_clusters, echo=FALSE, message=FALSE}

clusters_tf_idf <- clusters_tidy %>% 
  count(stage, word, sort = TRUE) %>% 
  bind_tf_idf(word, stage, n) %>%
  arrange(-tf_idf) %>%
  group_by(stage) %>% 
  top_n(9) %>% 
  ungroup 

clusters_tf_idf %>% 
  mutate(word = reorder_within(word, tf_idf, stage)) %>% #reordering by tf_idf
  ggplot(aes(word, tf_idf, fill = stage)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~stage, scales = "free", ncol = 2)+
  scale_x_reordered() +
  coord_flip() +
  theme(strip.text=element_text(size=11)) +
  labs(x = NULL, y = "tf-idf",
       title = "Highest tf-idf words in tweets during the different stages of Colorado Flood",
       subtitle = "Words importance within tweets sent during each stage")

```

<br>



####  **Topic Modeling**

Again, after playing with different numbers, I decided to train a topic model with 5 topics. From 6 on, topics started to look very similar (with the same bag of words). Here is a summary of the results after this process: 

```{r topic_modeling_clusters, echo=FALSE, message=FALSE}

clusters_dfm <- clusters_tidy %>%
  count(stage, word, sort = TRUE) %>% 
  cast_dfm(stage, word, n) # 4 documents (4 stages). Quanteda document frecuency. Special implementation for document term matrix

# Let's train oir topic model

topic_model_clusters <- stm(clusters_dfm, K = 5, init.type = "Spectral") # unsupervised 
summary(topic_model_clusters)
```


```{r beta_clusters, echo=FALSE, message=FALSE}
td_beta_clusters <- tidy(topic_model_clusters) # The beta matrix shows what are the words that contribute to each topic.

td_beta_clusters %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  ungroup %>% 
  mutate(topic = paste0("Topic ", topic),
         term = reorder(term, beta)) %>% #reordering by tf_idf
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~topic, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics") # To see which words contribute the most to each topic
```


```{r gamma_clusters, echo=FALSE, message=FALSE}
td_gamma_clusters <- tidy(topic_model_clusters, matrix = "gamma",                    
                 document_names = rownames(clusters_dfm))

ggplot(td_gamma_clusters, aes(gamma, fill = as.factor(topic))) +
  geom_histogram(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~ topic, ncol = 2) +
  labs(title = "Distribution of document probabilities for each topic",
       subtitle = "Each topic is associated with 1-3 stages",
       y = "Number of stages", x = expression(gamma))

```

```{r subset_by_county, echo=FALSE, message=FALSE}
library(tigris) # Loading tigris dataset/ Census
library(rgeos) #  rgeos required for finding out which hole belongs to which exterior ring

# Data to create basemap
counties <- counties("Colorado", cb = FALSE, resolution = "500k")
counties_sf <- st_as_sf(counties)
affected_counties <- counties_sf %>% 
  filter(NAME == "Arapahoe" | NAME == "Boulder" | NAME == "Denver"|
           NAME == "El Paso" | NAME ==  "Jefferson" | NAME ==  "Larimer" |
           NAME == "Logan" | NAME ==  "Morgan" | NAME ==  "Weld")

st_crs(affected_counties)

```

```{r subset_projection, echo=FALSE, message=FALSE}

affected_counties_p <- st_transform(affected_counties, 4326)

st_crs(affected_counties_p)


```


```{r point_in_polygon, echo=FALSE, message=FALSE}
tweet_in_county <- st_join(colo_sf_p, affected_counties_p) %>% 
  select(county_name = NAME,
         date = date,
         state = state,
         text = text,
         user = user,
         u_profile = u_profile,
         stage = stage,
         geometry = geometry) %>% 
  filter(!is.na(county_name))

```

```{r point_in_polygon_proj, echo=FALSE, message=FALSE}

tweet_in_county_p <- tweet_in_county %>% st_transform(crs ="+proj=lcc +lat_1=20 + lat_2=60 + lat_0=40 +
               lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +
               units=m no_defs") # Projected in North_America_Lambert_Conformal_Conic

st_crs(tweet_in_county_p)
```

```{r point_in_polygon_plot, echo=FALSE, message=FALSE, eval= FALSE}

TweetInCounty_Map <- leaflet(tweet_in_county_p) %>% # Interactive map to see resulting clusters
  addTiles()  %>%
  addProviderTiles(providers$OpenStreetMap.BlackAndWhite) %>% 
  addCircles(tweet_in_county_p,
             weight = 3, 
             radius=40,
             color= "Blue", 
             stroke = TRUE, 
             fillOpacity = 0.5)%>% 
  setView(lng = -105, lat = 40, zoom = 8.5)

TweetInCounty_Map

```



####  **Most common words (Within Spatial Clusters)**

A quick view of the most common words in the whole dataset: 

```{r data_subsets_tidy2, echo=FALSE, message=FALSE}

#This is mainly to remove geometry to make computation faster

tweet_in_county_df <- tweet_in_county_p %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  st_set_geometry(NULL)

```


```{r one-token-per-document-per-row-clusters2, echo=FALSE, message=FALSE}

tweet_in_county_tidy <- tweet_in_county_df %>% 
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words)  # remove stopwords
  
tweet_in_county_tidy %>% 
  count(word, sort = TRUE) # Count words

```


Again, since **"boulder"** is the most common word and is going to have a big effect in our topic modelling, it was removed from the dataset. The following four terms *("boulderflood", "colorado", "cowx", "flood")* were also excluded because they were so common and used neutrally in all four stages. After excluding these six terms, the new list of common words looks as follows:
                                 
                                 
```{r remove_words_clusters2, echo=FALSE, message=FALSE}

keywords_clusters <- c("boulder", "boulderflood", "colorado", "cowx", "flood", "coflood")

tweet_in_county_tidy <- tweet_in_county_df %>% 
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) %>%   # remove stopwords
  filter(!word %in% keywords_clusters) 

tweet_in_county_tidy %>% 
  count(word, sort = TRUE) # Count words

```


```{r explore_tf_idf_clusters2, echo=FALSE, message=FALSE}

tweet_in_county_tf_idf <- tweet_in_county_tidy %>% 
  count(stage, word, sort = TRUE) %>% 
  bind_tf_idf(word, stage, n) %>%
  arrange(-tf_idf) %>%
  group_by(stage) %>% 
  top_n(9) %>% 
  ungroup 

tweet_in_county_tf_idf %>% 
  mutate(word = reorder_within(word, tf_idf, stage)) %>% #reordering by tf_idf
  ggplot(aes(word, tf_idf, fill = stage)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~stage, scales = "free", ncol = 2)+
  scale_x_reordered() +
  coord_flip() +
  theme(strip.text=element_text(size=11)) +
  labs(x = NULL, y = "tf-idf",
       title = "Highest tf-idf words in tweets during the different stages of Colorado Flood",
       subtitle = "Words importance within tweets sent during each stage")

```

<br>



####  **Topic Modeling**

Again, after playing with different numbers, I decided to train a topic model with 5 topics. From 6 on, topics started to look very similar (with the same bag of words). Here is a summary of the results after this process: 

```{r topic_modeling_clusters2, echo=FALSE, message=FALSE}

tweet_in_county_dfm <- tweet_in_county_tidy %>%
  count(stage, word, sort = TRUE) %>% 
  cast_dfm(stage, word, n) # 4 documents (4 stages). Quanteda document frecuency. Special implementation for document term matrix

# Let's train oir topic model

topic_model_tweet_in_county <- stm(tweet_in_county_dfm, K = 4, init.type = "Spectral") # unsupervised 
summary(topic_model_tweet_in_county)
```


```{r beta_clusters2, echo=FALSE, message=FALSE}
td_beta_tweet_in_county <- tidy(topic_model_tweet_in_county) # The beta matrix shows what are the words that contribute to each topic.

td_beta_tweet_in_county %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  ungroup %>% 
  mutate(topic = paste0("Topic ", topic),
         term = reorder(term, beta)) %>% #reordering by tf_idf
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~topic, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics") # To see which words contribute the most to each topic
```


```{r gamma_clusters2, echo=FALSE, message=FALSE}
td_gamma_tweet_in_county <- tidy(topic_model_tweet_in_county, matrix = "gamma",                    
                 document_names = rownames(tweet_in_county_dfm))

ggplot(td_gamma_tweet_in_county, aes(gamma, fill = as.factor(topic))) +
  geom_histogram(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~ topic, ncol = 2) +
  labs(title = "Distribution of document probabilities for each topic",
       subtitle = "Each topic is associated with 1-3 stages",
       y = "Number of stages", x = expression(gamma))

```
