---
title: "Colorado"
author: "Marcela Suárez"
date: "July 18, 2019"
output: html_document
---
<br>


This document presents the results obtained for Colorado.  

<br>

#### **A general  overview of the dataset**

```{r import_data, echo=FALSE, message=FALSE}

library(tidyverse) 
colorado <- read_csv("ColoradoFloodOriginal.csv") # Loading data into R

 # Create a subset with the variables to be used
library(lubridate) # For dates
colorado <- colorado %>% 
  mutate(date = ymd_hms(colorado$created_at, tz='America/Denver')) %>% # Mountain Daylight Time. Time zone in Colorado (GMT-6)
  select(lat = latitude, 
         lon = longitude, 
         date = date, 
         state = c_state,
         text = t_text,
         user = u_name,
         u_profile = u_description)

str(colorado) # shows the structure of the dataframe
```

Tweets were collected from `r min(colorado$date)` to `r max(colorado$date)`  

<br>


####  **Temporal stages**

Four stages were defined:

```{r define_temporal_stages, echo=FALSE, message=FALSE}

colorado <- colorado %>%  # The way to define them in a tidyway
  mutate(stage = ifelse(date <= "2013-09-11 00:00:00 MDT", "preflood", 
                        ifelse(date <= "2013-09-16 00:00:00 MDT", "flood",
                               ifelse(date <= "2013-09-23 00:00:00 MDT", "immediate_aftermath", "postflood")))) %>%
    mutate(stage = factor(stage, levels = unique(stage))) # Turning stages into factors to use them for plotting later
```

* *Preflood*: From `r min(colorado$date)` to 2013-09-11 00:00:00.
* *Flood*: From 2013-09-11 00:00:00 to 2013-09-16 00:00:00.
* *Immediate Aftermath*: From 2013-09-16 00:00:00 to 2013-09-23 00:00:00.
* *Postflood*: From 2013-09-23 00:00:00 to `r max(colorado$date)`.

<br>


####  **Cleaning dataset**

* *Removing retweets:*
```{r remove_retweets, echo=FALSE, message=FALSE}

library(tidytext)
library(stringr)

colorado <- colorado %>% 
  filter(!str_detect(text, "^RT"))# There will remain some retuits since those will be commented retweets (actually adding some info)
```
    From the original dataset containing 5915 tweets, only 5846 were retained.  

* *Removing tweets sent by bots:*
```{r remove_bots, echo=FALSE, message=FALSE}
colorado <- colorado %>% 
  filter(!str_detect(u_profile, "TwitZip")) 
```
    Tweets sent by [TwitZIP](http://twitzip.com/), a bot service to discover local neighborhood news, were removed. After this process 4916 were retained. 
    
* *Removing urls, usernames, punctuation and numbers:*
```{r clean_tweets, echo=FALSE, message=FALSE}

colo_clean <- colorado %>% 
   mutate(text = str_remove_all(text, 'http[^ ]+'), #Finally removing urls in a nice way
          text = str_remove_all(text, '@[^ ]+'), # Removing usernames
          text = str_remove_all(text, "[:punct:]"), # Removing puntuation
          text = str_remove_all(text, "[:digit:]")) # Removing numbers
```
    All references to usernames, urls and punctuation and numbers were removed in order to prepare the dataset for tokenization.
<br>



####  **Most common words**

A quick view of the most common words in the whole dataset: 

```{r one-token-per-document-per-row, echo=FALSE, message=FALSE}

colo_tidy <- colo_clean %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) # remove stopwords

colo_tidy %>% 
  count(word, sort = TRUE) # Count words
```

Since **boulder** is the most common word and is going ot have a big effect in our topic modelling, it was removed from the dataset.
                                 
                                 
```{r remove_words, echo=FALSE, message=FALSE}

keywords <- c("boulder", "boulderflood", "coflood", "cofloodrelief")
colo_tidy <- colo_clean %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) %>%  # remove stopwords
  filter(!word %in% keywords) # Next two candidates to be removed are Im and boulderflood. TO CHECK LATER!

colo_tidy %>% 
  count(word, sort = TRUE) # Count words
```

<br>

tf_idf was explored in order to identify which words are important in each of the flood stages.

```{r explore_tf_idf, echo=FALSE, message=FALSE}

colo_tf_idf <- colo_tidy %>% 
  count(stage, word, sort = TRUE) %>% 
  bind_tf_idf(word, stage, n) %>% 
  group_by(stage) %>% 
  top_n(9) %>% 
  ungroup 

colo_tf_idf %>% 
  mutate(word = reorder(word, tf_idf)) %>% 
  ggplot(aes(word, tf_idf, fill = stage)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~stage, scales = "free") +
  coord_flip()
  
```


