---
title: "Colorado"
author: "Marcela Suárez"
date: "November 19, 2019"
output: html_document
---
<br>


This document presents the results obtained for Colorado.  

<br>

## **Analysis with the whole dataset**

#### **A general  overview of the dataset**

```{r import_data, echo=FALSE, message=FALSE, tidy=TRUE}

library(tidyverse) 
colorado <- read_csv("ColoradoFloodOriginal.csv") # Loading data into R

 # Create a subset with the variables to be used
library(lubridate) # For dates
colorado <- colorado %>% 
  mutate(date = ymd_hms(colorado$created_at, tz='America/Denver')) %>% # Mountain Daylight Time. Time zone in Colorado (GMT-6)
  select(lat = latitude, 
         lon = longitude, 
         date = date, 
         state = c_state,
         text = t_text,
         user = u_name,
         u_profile = u_description)

str(colorado) # shows the structure of the dataframe
```

Tweets were collected from `r min(colorado$date)` to `r max(colorado$date)`  

<br>


####  **Cleaning dataset**

* *Removing retweets:*
```{r remove_retweets, echo=FALSE, message=FALSE}

library(tidytext)
library(stringr)

colorado <- colorado %>% 
  filter(!str_detect(text, "^RT")) %>% 
  filter(!str_detect(text, "#TweetMyJobs")) %>% 
  filter(!str_detect(text, "#GJCO")) %>% 
  filter(!str_detect(text, "#Job")) %>% 
  filter(!str_detect(text, "I'm")) # There will remain some retuits since those will be commented retweets (actually adding some info)
```
    From the original dataset containing 5915 tweets, only 4837 were retained.  

* *Removing tweets sent by bots:*
```{r remove_bots, echo=FALSE, message=FALSE}
colorado <- colorado %>% 
  filter(!str_detect(u_profile, "TwitZip")) %>% 
  filter(!str_detect(u_profile, "TMJ-COB PM Jobs")) %>% 
  filter(!str_detect(u_profile, "JobsDirectUSA.com")) %>% 
  filter(!str_detect(u_profile, "Whole Foods")) %>% 
  filter(!str_detect(u_profile, "Superpages Colorado"))

```
    Tweets sent by [TwitZIP](http://twitzip.com/), a bot service to discover local neighborhood news, were removed. After this process 4086 were retained. Other three bot accounts *("TMJ-COB PM Jobs", "JobsDirectUSA.com", "Whole Foods")* were detected and tweets were removed. A total of 4086 tweets were retained.
    
* *Removing urls, usernames, punctuation and numbers:*
```{r clean_tweets, echo=FALSE, message=FALSE}

colo_clean <- colorado %>% 
   mutate(text = str_remove_all(text, 'http[^ ]+'), #Finally removing urls in a nice way
          text = str_remove_all(text, '@[^ ]+'), # Removing usernames
          text = str_remove_all(text, "[:punct:]"), # Removing puntuation
          text = str_remove_all(text, "[:digit:]"), # Removing numbers
          text = str_remove_all(text, "|")) 
```
    All references to usernames, urls and punctuation and numbers were removed from each tweet in order to prepare the dataset for tokenization.



```{r tweets distribution, echo=FALSE, message=FALSE, warning=FALSE}
library(scales)
gen_hist <- ggplot(colo_clean, aes(x = colo_clean$date)) +
  geom_histogram(aes(y = ..density..), colour = "#878787", fill = "white", binwidth = 3600) +
  geom_density(alpha=.2, colour = "#35978f", fill="#35978f") +# Overlay with transparent density plot
  scale_x_datetime(name = "Date", breaks = date_breaks("2 day"),
                   labels = date_format("%m/%d"), expand = c(0.01,0),
                   limits = c(
                     as.POSIXct(min(colo_clean$date)),
                     as.POSIXct(max(colo_clean$date))
                   )) + scale_y_continuous(name = "Density") +
  ggtitle("")   

gen_hist

```


```{r days_asfactor, echo=FALSE, message=FALSE}

colo_clean <- colo_clean %>% 
  mutate(day = as.factor(day(date))) %>%
  mutate(tweet_id = row_number())


sep15_count <- colo_clean %>% 
  filter(day == 15) %>% 
  nrow()
```

Note: There were only 5 observations on September 15 and only 131 on September 16.

In this stage almost **6%** of the tweets were excluded. 228 tweets from the 4086 total tweets.

```{r remove_topuser, echo=FALSE, message=FALSE}

user_frequency <- colo_clean %>%
        group_by(user) %>%
        count()

colo_clean <- colo_clean %>% 
  filter(user != "Denver CP")
```



<br>


####  **Temporal stages**

Four stages were defined:

```{r define_temporal_stages, echo=FALSE, message=FALSE}

colo_clean <- colo_clean %>%  # The way to define them in a tidyway
  mutate(stage = ifelse(date <= "2013-09-11 00:00:00 MDT", "preflood", 
                        ifelse(date <= "2013-09-16 00:00:00 MDT", "flood",
                               ifelse(date <= "2013-09-23 00:00:00 MDT", "immediate_aftermath", "postflood")))) %>%
  mutate(stage = factor(stage, levels = unique(stage))) %>% # Turning stages into factors to use them for plotting later
  mutate(stage = factor(stage,levels(stage)[c(3, 4, 2, 1)])) # To plot in an specific order

```

* *Preflood*: From `r min(colorado$date)` to 2013-09-11 00:00:00.
* *Flood*: From 2013-09-11 00:00:00 to 2013-09-16 00:00:00.
* *Immediate Aftermath*: From 2013-09-16 00:00:00 to 2013-09-23 00:00:00.
* *Postflood*: From 2013-09-23 00:00:00 to `r max(colorado$date)`.

Total number of tweets by stage:

```{r tweets_by_stage, echo=FALSE, message=FALSE}
frequency_by_stage <- colo_clean %>%
        group_by(stage) %>%
        count()

knitr::kable(frequency_by_stage, format="html")

```


<br>



####  **Most common words**

A quick view of the most common words in the whole dataset: 

```{r one-token-per-document-per-row, echo=FALSE, message=FALSE}

colo_tidy <- colo_clean %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) # remove stopwords

colo_tidy %>% 
  count(word, sort = TRUE) # Count words
```

Since **"boulder"** is the most common word and is going ot have a big effect in our topic modelling, it was removed from the dataset. After this process the new list of common words looks as follows:
                                 
                                 
```{r remove_words, echo=FALSE, message=FALSE}

keywords <- c("boulder", "boulderflood", "cowx", "$", "amp", "|", "rt", "coflood", "colorado")

colo_tidy <- colo_clean %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) %>%  # remove stopwords
  filter(!word %in% keywords) # Next two candidates to be removed are Im and boulderflood. TO CHECK LATER!

colo_tidy %>% 
  count(word, sort = TRUE) # Count words
```


```{r save_files, echo=FALSE, message=FALSE}
saveRDS(colo_clean, file = "colo_clean.rds")
saveRDS(colo_tidy, file = "colo_tidy.rds")
```


<br>

The statistic **tf-idf** was computed in order to identify which words are important in each of the flood stages. It measures how important a word is to a document in a collection (or corpus) of documents, in our case, it measures how important a word is to a tweet in a collection of tweets. In this case the collection of tweets was the set of tweets belonging to each stage.

```{r explore_tf_idf, echo=FALSE, message=FALSE, fig.width=9, fig.height=8}

colo_tf_idf <- colo_tidy %>% 
  count(stage, word, sort = TRUE) %>% 
  bind_tf_idf(word, stage, n) %>%
  arrange(-tf_idf) %>%
  group_by(stage) %>% 
  top_n(10) %>% 
  ungroup 

colo_tf_idf %>% 
  mutate(word = reorder_within(word, tf_idf, stage)) %>% #reordering by tf_idf
  ggplot(aes(word, tf_idf, fill = stage)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~stage, scales = "free", ncol = 2)+
  scale_x_reordered() +
  coord_flip() +
  theme(strip.text=element_text(size=11)) +
  labs(x = NULL, y = "tf-idf",
       title = "Highest tf-idf words in tweets during the different stages of Colorado Flood",
       subtitle = "Words importance within tweets sent during each stage")


```

<br>



####  **Topic Modeling**


```{r create_matrix, echo=FALSE, message=FALSE}
library(stm) #fast compared to other implementations of topic models. Base don C++
library(quanteda)

# Either one (of the following) works as input for the topic modelling

# Create Document-Term Matrix
colo_dfm <- colo_tidy %>%
  count(tweet, word, sort = TRUE) %>% 
  cast_dfm(tweet, word, n) # 4 documents (4 stages). Quanteda document frecuency. Special implementation for document term matrix

# Create Sparce Matrix
colo_sparse <- colo_tidy %>%
  count(tweet_id, word) %>%
  cast_sparse(tweet_id, word, n)

```


```{r plot_topic_modeling_results, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}

k_result <- readRDS(file = "k_result.rds")

k_result %>%
  transmute(K,
            `Lower bound` = lbound,
            Residuals = map_dbl(residual, "dispersion"),
            `Semantic coherence` = map_dbl(semantic_coherence, mean),
            `Held-out likelihood` = map_dbl(eval_heldout, "expected.heldout")) %>%
  gather(Metric, Value, -K) %>%
  ggplot(aes(K, Value, color = Metric)) +
  geom_line(size = 1.5, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~Metric, scales = "free_y") +
  labs(x = "K (number of topics)",
       y = NULL,
       title = "Model diagnostics by number of topics",
       subtitle = "These diagnostics indicate that a good number of topics would be around 13")
```

```{r plot_semantic_and_exclusivity, echo=FALSE, message=FALSE, fig.width=9, fig.height=6}

k_result %>%
  select(K, exclusivity, semantic_coherence) %>%
  filter(K %in% c(11,17,22)) %>% # This values should be the same used when creating the model in the R script
  unnest() %>%
  mutate(K = as.factor(K)) %>%
  ggplot(aes(semantic_coherence, exclusivity, color = K)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(x = "Semantic coherence",
       y = "Exclusivity",
       title = "Comparing exclusivity and semantic coherence",
       subtitle = "Models with fewer topics have higher semantic coherence for more topics, but lower exclusivity")

```

```{r beta, echo=FALSE, message=FALSE, fig.width=10, fig.height=8}
topic_model <- k_result %>% 
  filter(K == 17) %>% 
  pull(topic_model) %>% 
  .[[1]]

# topic_model

td_beta <- tidy(topic_model) # The beta matrix shows what are the words that contribute to each topic.

td_beta %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  ungroup %>% 
  mutate(topic = paste0("Topic ", topic),
         term = reorder(term, beta)) %>% #reordering by tf_idf
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~topic, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics") # To see which words contribute the most to each topic
```


```{r gamma, echo=FALSE, message=FALSE, warning=FALSE}
library(ggthemes)
library(extrafont)

td_gamma <- tidy(topic_model, matrix = "gamma",                    
                 document_names = rownames(colo_sparse))

top_terms <- td_beta %>%
  arrange(beta) %>%
  group_by(topic) %>%
  top_n(7, beta) %>%
  arrange(-beta) %>%
  select(topic, term) %>%
  summarise(terms = list(term)) %>%
  mutate(terms = map(terms, paste, collapse = ", ")) %>% 
  unnest()

gamma_terms <- td_gamma %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma)) %>%
  arrange(desc(gamma)) %>%
  left_join(top_terms, by = "topic") %>%
  mutate(topic = paste0("Topic ", topic),
         topic = reorder(topic, gamma))

gamma_terms %>%
  top_n(17, gamma) %>%
  ggplot(aes(topic, gamma, label = terms, fill = topic, alpha = 0.7)) +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, nudge_y = 0.0005, size = 3,
            family = "CM Roman") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 0.30),
                     labels = percent_format()) +
  theme_tufte(base_family = "CM Roman",ticks = FALSE) +
  theme(plot.title = element_text(size = 14,
                                  family="CM Roman"),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  labs(x = NULL, y = expression(gamma),
       title = "Topics by prevalence in the dataset containing all Colorado tweets",
       subtitle = "With the top words that contribute to each topic")

```

```{r prepare_heatmap, echo=FALSE, message=FALSE}

heat <- td_gamma %>% 
  rename(tweetid = document) %>% 
  mutate(tweet_id = as.integer(tweetid)) %>% 
  left_join(colo_clean, by = "tweet_id") %>% 
  select(tweet_id, topic, gamma, day)

# Assigning a topic to each tweet based on which one was had the maximum Gamma
heat <- heat %>% 
  group_by(tweet_id) %>% 
  filter(gamma == max(gamma)) %>% 
  ungroup()

# Compute the total number of tweets belonging to each topic each day
heat <- heat %>% 
  group_by(day, topic) %>% 
  add_count() %>% 
  rename(total_topic_day = n) %>% 
  ungroup()

# Compute the total number of tweets per day
heat <- heat %>% 
  group_by(day) %>% 
  add_count() %>% 
  rename(total_tweets_day = n) %>% 
  mutate(dens = total_topic_day/total_tweets_day)

# Standarazing tweet density (total_topic_day). Computing Z-score values for total_topic_day
scale_this <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

scaled_heat <- 
  heat %>%
  group_by(day) %>%
  mutate(scaled_total_topic_day = scale_this(total_topic_day))

```


```{r gammaclusters2_heatplot, echo=FALSE, message=FALSE}
#Assign color variables
col1 = "#d8e1cf" 
col2 = "#438484"


library(RColorBrewer)
ggplot(scaled_heat, aes(day, as.factor(topic))) +
  geom_tile(aes(fill = scaled_total_topic_day), color = "white") +
  scale_fill_gradient(low = col1, high = col2) +  
  ylab("Topic") +
  xlab("Flood day") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=12,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "")
```



