---
title: "Flood Phase in Counties Warned by the NWS"
author: "Marcela"
date: "June 22, 2020"
output: html_document
---

<br>

####  **Subsetting by Affected Counties and Flood Phase**

**3858 tweets** sent during the disaster were filtered based on the 9 counties affected in the disaster --according to the reports: Arapahoe, Boulder, Denver, El Paso, Jefferson, Larimer, Logan, Morgan and Weld.

```{r subset_by_county, echo=FALSE, message=FALSE}
library(sf)
library(tidyverse)
library(tidytext)
library(scales)

# library(rgeos) #  rgeos required for finding out which hole belongs to which exterior ring

warned_counties <- readRDS(file = "Data/warned_counties.rds")
# st_crs(affected_counties)

colo_clean <- readRDS(file = "Data/colo_clean.rds")

```

```{r subset_projection, echo=FALSE, message=FALSE}

warned_counties_p <- st_transform(warned_counties, 4326)
# st_crs(affected_counties_p)

```


```{r getting_spatial, echo=FALSE, message=FALSE}

library(sf)

# Store tweets as simple features and project data
colo_sf_p <- colo_clean %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

# st_crs(colo_sf_p) # Retrieve current coord ref system: EPSG: 4326 WGS84

```


```{r point_in_polygon, echo=FALSE, message=FALSE}
flood_county_tweet <- st_join(colo_sf_p, warned_counties_p) %>% 
  select(county_name = NAME,
         date = date,
         state = state,
         text = text,
         user = user,
         u_profile = u_profile,
         stage = stage,
         day = day,
         tweet_id = tweet_id,
         geometry = geometry) %>% 
  filter(!is.na(county_name)) %>% 
  filter(stage == "flood")

```

Only **817** tweets sent during the "Flood" stage and from these counties were filtered for content analysis.

```{r create_map, echo=FALSE, message=FALSE, fig.width=9, fig.height=9}
library(mapview)

mapview(warned_counties_p, zcol = "NAME") +
  mapview(flood_county_tweet, 
          lwd = 0.5, 
          cex = 2, 
          popup = leafpop::popupTable(flood_county_tweet,
                             zcol = "text"))
```


#####  **Most common words (Within Affected Counties)**

A quick view of the most common words in the whole dataset: 

```{r data_subsets_tidy2, echo=FALSE, message=FALSE}

#This is mainly to remove geometry to make computation faster

flood_county_tweet_df <- flood_county_tweet %>% 
  mutate(tweet = text) %>% # To keep a column with the original tweet
  st_set_geometry(NULL)

```


```{r one-token-per-document-per-row-clusters2, echo=FALSE, message=FALSE}

flood_county_tweet_tidy <- flood_county_tweet_df %>% 
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words)  # remove stopwords
  
flood_county_tweet_tidy %>% 
  count(word, sort = TRUE) # Count words

```


Again, since **"boulder"** is the most common word and is going to have a big effect in our topic modelling, it was removed from the dataset. The following four terms *("boulder", "boulderflood", "cowx", "amp", "rt", "coflood")* were also excluded because they were so common and used neutrally in all four stages. After excluding these six terms, the new list of common words looks as follows:
                                 
                                 
```{r remove_words_clusters2, echo=FALSE, message=FALSE}

keywords <- c("boulder", "boulderflood", "cowx")

flood_county_tweet_tidy <- flood_county_tweet_df %>% 
  unnest_tokens(word, text, token = "tweets") %>% # add a row for each token (word) and repeat the other information
  anti_join(stop_words) %>%   # remove stopwords
  filter(!word %in% keywords) 

flood_county_tweet_tidy %>% 
  count(word, sort = TRUE) # Count words

```


```{r explore_tf_idf_clusters2, echo=FALSE, message=FALSE, fig.width=9, fig.height=8}

flood_county_tweet_tf_idf <- flood_county_tweet_tidy %>% 
  count(day, word, sort = TRUE) %>% 
  bind_tf_idf(word, day, n) %>%
  arrange(-tf_idf) %>%
  group_by(day) %>% 
  top_n(10) %>% 
  ungroup 

flood_county_tweet_tf_idf %>% 
  mutate(word = reorder_within(word, tf_idf, day)) %>% #reordering by tf_idf
  ggplot(aes(word, tf_idf, fill = day)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~day, scales = "free", ncol = 2)+
  scale_x_reordered() +
  coord_flip() +
  theme(strip.text=element_text(size=11)) +
  labs(x = NULL, y = "tf-idf",
       title = "Highest tf-idf words in tweets during the different stages of Colorado Flood",
       subtitle = "Words importance within tweets sent during each stage")

```

<br>



#####  **Topic Modeling**


```{r create_matrix, echo=FALSE, message=FALSE}

library(stm) #fast compared to other implementations of topic models. Base don C++
library(quanteda)

# Either one (of the following) works as input for the topic modelling

# Create Document-Term Matrix
flood_county_tweet_dfm <- flood_county_tweet_tidy %>%
  count(tweet, word, sort = TRUE) %>% 
  cast_dfm(tweet, word, n) # 4 documents (4 stages). Quanteda document frecuency. Special implementation for document term matrix

# Create Sparce Matrix
flood_county_tweet_sparse <- flood_county_tweet_tidy %>%
  count(tweet_id, word) %>%
  cast_sparse(tweet_id, word, n)

saveRDS(flood_county_tweet_sparse, file = "Data/flood_warnedcounties_tweet_sparse.rds")
```




```{r plot_topic_modeling_results, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}

k_result_flood_county <- readRDS(file = "Data/k_result_warned_counties.rds")

k_result_flood_county %>%
  transmute(K,
            `Lower bound` = lbound,
            Residuals = map_dbl(residual, "dispersion"),
            `Semantic coherence` = map_dbl(semantic_coherence, mean),
            `Held-out likelihood` = map_dbl(eval_heldout, "expected.heldout")) %>%
  gather(Metric, Value, -K) %>%
  ggplot(aes(K, Value, color = Metric)) +
  geom_line(size = 1.5, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~Metric, scales = "free_y") +
  labs(x = "K (number of topics)",
       y = NULL,
       title = "Model diagnostics by number of topics",
       subtitle = "These diagnostics indicate that a good number of topics would be around 13")
```


```{r plot_semantic_and_exclusivity, echo=FALSE, message=FALSE, fig.width=9, fig.height=6}

k_result_flood_county %>%
  select(K, exclusivity, semantic_coherence) %>%
  filter(K %in% c(6,9)) %>% # This values should be the same used when creating the model in the R script
  unnest() %>%
  mutate(K = as.factor(K)) %>%
  ggplot(aes(semantic_coherence, exclusivity, color = K)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(x = "Semantic coherence",
       y = "Exclusivity",
       title = "Comparing exclusivity and semantic coherence",
       subtitle = "Models with fewer topics have higher semantic coherence for more topics, but lower exclusivity")

```


```{r beta, echo=FALSE, message=FALSE, fig.width=13, fig.height=8}
topic_model_flood_county <- k_result_flood_county %>% 
  filter(K == 9) %>% 
  pull(topic_model) %>% 
  .[[1]]

# topic_model

td_beta_flood_county <- tidy(topic_model_flood_county) # The beta matrix shows what are the words that contribute to each topic.

td_beta_flood_county %>% 
  group_by(topic) %>% 
  top_n(10, beta) %>% 
  ungroup %>% 
  mutate(topic = paste0("Topic ", topic),
         term = reorder(term, beta)) %>% #reordering by tf_idf
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~topic, scales = "free_y") +
  coord_flip() +
  scale_x_reordered() +
  labs(x = NULL, y = expression(beta),
       title = "Highest word probabilities for each topic",
       subtitle = "Different words are associated with different topics") # To see which words contribute the most to each topic
```



```{r gamma, echo=FALSE, message=FALSE, warning=FALSE, width = 12}
library(ggthemes)
library(extrafont)

td_gamma_flood_county <- tidy(topic_model_flood_county, matrix = "gamma",                    
                 document_names = rownames(flood_county_tweet_sparse))

top_terms_flood_county <- td_beta_flood_county %>%
  arrange(beta) %>%
  group_by(topic) %>%
  top_n(12, beta) %>%
  arrange(-beta) %>%
  select(topic, term) %>%
  summarise(terms = list(term)) %>%
  mutate(terms = map(terms, paste, collapse = ", ")) %>% 
  unnest()

gamma_terms_flood_county <- td_gamma_flood_county %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma)) %>%
  arrange(desc(gamma)) %>%
  left_join(top_terms_flood_county, by = "topic") %>%
  mutate(topic = paste0("Topic ", topic),
         topic = reorder(topic, gamma))

gamma_terms_flood_county %>%
  top_n(17, gamma) %>%
  ggplot(aes(topic, gamma, label = terms, fill = topic, alpha = 0.7)) +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, nudge_y = 0.0005, size = 3,
            family = "CM Roman") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 0.30),
                     labels = percent_format()) +
  theme_tufte(base_family = "CM Roman",ticks = FALSE) +
  theme(plot.title = element_text(size = 14,
                                  family="CM Roman"),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  labs(x = NULL, y = expression(gamma),
       title = "Topics by prevalence in the dataset containing all Colorado tweets",
       subtitle = "With the top words that contribute to each topic")

```

```{r prepare_heatmap, echo=FALSE, message=FALSE}

heat_flood_county <- td_gamma_flood_county %>% 
  rename(tweetid = document) %>% 
  mutate(tweet_id = as.integer(tweetid)) %>% 
  left_join(colo_clean, by = "tweet_id") %>% 
  select(tweet_id, topic, gamma, day)

# Assigning a topic to each tweet based on which one was had the maximum Gamma
heat_flood_county <- heat_flood_county %>% 
  group_by(tweet_id) %>% 
  filter(gamma == max(gamma)) %>% 
  ungroup()

# Compute the total number of tweets belonging to each topic each day
heat_flood_county <- heat_flood_county %>% 
  group_by(day, topic) %>% 
  add_count() %>% 
  rename(total_topic_day = n) %>% 
  ungroup()

# Compute the total number of tweets per day
heat_flood_county <- heat_flood_county %>% 
  group_by(day) %>% 
  add_count() %>% 
  rename(total_tweets_day = n) %>% 
  mutate(dens = total_topic_day/total_tweets_day)

# Standarazing tweet density (total_topic_day). Computing Z-score values for total_topic_day
scale_this <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

scaled_heat_flood_county <- 
  heat_flood_county %>%
  group_by(day) %>%
  mutate(scaled_total_topic_day = scale_this(total_topic_day))

```


```{r gammaclusters2_heatplot, echo=FALSE, message=FALSE}
#Assign color variables
col1 = "#d8e1cf" 
col2 = "#438484"


library(RColorBrewer)
ggplot(scaled_heat_flood_county, aes(day, as.factor(topic))) +
  geom_tile(aes(fill = scaled_total_topic_day), color = "white") +
  scale_fill_gradient(low = col1, high = col2) +  
  ylab("Topic") +
  xlab("Flood day") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=12,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(fill = "")
```


```{r joinTweetsTopics, echo=FALSE, message=FALSE}

tweet_and_topic <- dplyr::left_join(heat_flood_county, colo_clean, by = "tweet_id") %>% 
  select(tweet_id = tweet_id,
         topic = topic,
         text, text,
         user = user,
         u_profile = u_profile,
         date = date,
         lat = lat,
         lon = lon,
         day = day.x) %>% 
  mutate(topic = paste0("Topic ", topic))

tweet_and_topic$topic <- factor(tweet_and_topic$topic)
```

```{r geoconversion, echo=FALSE, message=FALSE}
tweet_and_topic_geo <- tweet_and_topic %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)
```

```{r topicMap, echo=FALSE, message=FALSE}
library(colorspace)
library(RColorBrewer)
library(leaflet)
library(htmltools)


markerCol <- colorFactor(palette = brewer.pal(12, "Paired"), tweet_and_topic_geo$topic)

TopicalMap <- leaflet(tweet_and_topic_geo) %>% # Interactive map to see resulting clusters
  addTiles()  %>%
  addProviderTiles(providers$OpenStreetMap.BlackAndWhite) %>% 
  addCircles(weight = 3, 
             radius=40,
             color= markerCol(tweet_and_topic_geo$topic), 
             stroke = TRUE, 
             fillOpacity = 0.5,
             popup = ~htmlEscape(tweet_and_topic_geo$text))%>% 
  setView(lng = -105, lat = 40, zoom = 8.5)

TopicalMap

htmlwidgets::saveWidget(TopicalMap, file = "MapTopicsFloodCounties.html")

```



```{r computing_distance, echo=FALSE, message=FALSE}
damagePols <- st_read("Data/GroundTruthData/StructuresPols/DamagePolsAllIncluded.geojson")


# Check if there are points within the polygon. St_within also works for this
pnts <- tweet_and_topic_geo %>% 
  mutate(intersection = as.integer(st_intersects(geometry, damagePols))) 

# Compute distances from points to the closest polygon
dist <- geosphere::dist2Line(p = st_coordinates(pnts), line = st_coordinates(damagePols)[,1:2])

# bind results with original points
pts.wit.dist <- cbind(pnts, dist) %>% 
  mutate(distancee = ifelse(is.na(intersection), distance, 0)) %>% 
  select(-c(intersection, distance))


```

```{r plotting_distance, echo=FALSE, message=FALSE, fig.width=10, fig.height=10}

pts.wit.dist %>% 
  ggplot(aes(distancee, date, color = topic)) + # 
  geom_point(size = 2, alpha = 0.7) +
  labs(x = "Distance",
       y = "Date",
       title = "Distance between tweets and nearest reported structure",
       subtitle = "Distance to the nearest reported structure and tweet's date") +
  facet_wrap(~topic) +
  theme(legend.position="bottom")
```

```{r distance_histo, echo=FALSE, message=FALSE}
ggplot(pts.wit.dist, aes(x = distancee)) +
  geom_density(alpha=.2, colour = "#35978f", fill="#35978f")
```


```{r b,fig.width=14, fig.height=10}

library(viridis)
library(hrbrthemes)

pts.wit.dist2 <- pts.wit.dist %>% 
  filter(distancee <= 12500) 

pts.wit.dist2 %>% 
  ggplot(aes(distancee, color = topic)) + 
  geom_freqpoly(binwidth = 500)+
  xlab("Distance") +
  ylab("Number of tweets by distance band") +
  scale_x_continuous(breaks = seq(0, 13000, 1000), expand = c(0, 0), limits = c(0,NA)) +
  scale_color_viridis(discrete=TRUE, guide=FALSE) +
  theme(legend.title = element_blank(),
        legend.background = element_rect(size = 2),
        legend.text = element_text(size = 10),
        plot.background = element_rect(colour = NA),
        axis.title = element_text(size = 12, face = "bold",
                                  margin=margin(30,20,0,0)),
        axis.line = element_line(colour = "Black"),
        axis.text = element_text(size = 7, angle = 90, color = "black"),
        strip.text = element_text(size = 11),
        legend.key.width = unit(1.5, "cm")) +
    facet_wrap(~topic) +
  geom_freqpoly(data=transform(pts.wit.dist2, topic=NULL),
                aes(y = ..count.. / n_distinct(pts.wit.dist$topic)), 
                binwidth = 500, 
                color = "red",
                alpha = 0.5,
                linetype = "longdash")

```



```{r b,fig.width=14, fig.height=10}

library(viridis)
library(hrbrthemes)

pts.wit.dist %>% 
  filter(distancee <= 12500) %>% 
  ggplot(aes(distancee, color = topic)) + 
  geom_freqpoly(binwidth = 500) +
  geom_freqpoly(aes(y = ..count.. / n_distinct(pts.wit.dist$topic)), 
                binwidth = 500, 
                color = "red",
                linetype = "longdash") +
  xlab("Distance") +
  ylab("Number of tweets by distance band") +
  scale_x_continuous(breaks = seq(0, 13000, 1000), expand = c(0, 0), limits = c(0,NA)) +
  scale_color_viridis(discrete=TRUE) +
  theme(legend.title = element_blank(),
        legend.background = element_rect(size = 2),
        legend.text = element_text(size = 10),
        plot.background = element_rect(colour = NA),
        axis.title = element_text(size = 12, face = "bold",
                                  margin=margin(30,20,0,0)),
        axis.line = element_line(colour = "Black"),
        axis.text = element_text(size = 7, angle = 90, color = "black"),
        strip.text = element_text(size = 11),
        legend.key.width = unit(1.5, "cm")) 

```